Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-07 main.tex
CP

Conventions:
%
\newcommand{\word}{W}
\newcommand{\addr}{A}
\newcommand{\WORDS}{{\mathcal W}}
\newcommand{\reg}{R}
\newcommand{\REGS}{{\mathcal R}}
\newcommand{\mach}{M}
\newcommand{\MACHS}{{\mathcal M}}
\newcommand{\PC}[1]{\mathit{PC}(#1)}
%
\begin{itemize}
\item Words / addresses: $\word,\addr \in \WORDS$
\item Registers: $\reg \in \REGS$
\item Machine states:
$\mach \in \MACHS = (\WORDS\rightarrow\WORDS,\, \REGS\rightarrow\WORDS,\,
\WORDS)$
\item Abbreviations: \\
$\mach(\addr) = \mach.1(\addr)$ \\
$\mach(\reg) = \mach.2(\reg)$ \\
$\mach(\textsc{pc}) = \mach.3$ \\
\end{itemize}

@@

\begin{itemize}
\item We use the word {\em policy} for any kind of dynamic enforcement
mechanism (e.g., a runtime taint tracker, heap coloring rules, capability
hardware, etc.). We use {\em property} for a mathematical predicate on
programs (or machine states) capturing some useful security goal, e.g. noninterference.

\item Reminder of the standard definition of NI.

@@

>>>>

Conventions:
%
\newcommand{\word}{W}
\newcommand{\addr}{A}
\newcommand{\WORDS}{{\mathcal W}}
\newcommand{\reg}{R}
\newcommand{\REGS}{{\mathcal R}}
\newcommand{\mach}{M}
\newcommand{\MACHS}{{\mathcal M}}
\newcommand{\PC}[1]{\mathit{PC}(#1)}
\newcommand{\pol}{P}
\newcommand{\POLS}{{\mathcal P}}
\newcommand{\trace}[1]{\overline{#1}}
\newcommand{\TRACES}{{\mathcal T}}
\newcommand{\stepsto}{\Rightarrow}
%
\begin{itemize}
\item Words / addresses: $\word,\addr \in \WORDS$
\item Registers: $\reg \in \REGS$
\item Machine states:
$\mach \in \MACHS = (\WORDS\rightarrow\WORDS,\, \REGS\rightarrow\WORDS,\,
\WORDS)$
\item Abbreviations:
\\ $\mach(\addr) = \mach.1(\addr)$
\\ $\mach(\reg) = \mach.2(\reg)$
\\ $\mach(\textsc{pc}) = \mach.3$
\item Step function\footnote{\bcp{I'm taking the step function to be total;
    we could make it partial if we want, but I'm not sure this refinement is
    important.}}: $\mathord{\stepsto} \in \MACHS \rightarrow \MACHS$
\end{itemize}

A {\em trace} $\trace{M} \in \TRACES$ is a nonempty, finite or infinite
sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
$\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.

We use the word {\em policy} for any kind of dynamic enforcement
mechanism (e.g., a runtime taint tracker, heap coloring rules, capability
hardware, etc.). We use {\em property} for a mathematical predicate on
programs (or machine states) capturing some useful security goal, e.g. noninterference.

Formally, a {\em policy} $\pol \in \POLS$ is a prefix-closed set of traces.
I.e., it is a restriction of the set of all traces generated by the standard
step function, where some traces may be ``truncated early.''

\iftext \bcp{Not sure we need names for $|TRACES$ or $\POLS$.} \fi

\bcp{Now we need to define {\em property} formally.  But since we're talking
about hyperproperties, I'm not certain how to go about it.  The standard way
would be something like ``A policy $P$ has property $Q$ (i.e., $P$ is
noninterfering) if, for every pair of traces $\trace{M}$ and $\trace{M'}$
in $P \times P$, if the initial states $M_0$ and $M'_0$ differ only in
high parts then every pair ...'' (blah: I'm not sure whether to talk about
lockstep-corresponding states, which seems to restrictive, or final states,
which doesn't work if the traces are infinite!  do we really need some kind
of reactive noninterference here?? Urg.)}

@@

\begin{itemize}
\item Reminder of the standard definition of NI.

@@

Presentation idea: Start here with a super-simple separation policy, which
just makes some part of the address space inaccessible, truncating any
traces that read or write it.  Formulate the property as a flavor of NI.
(The $+$ part in the title is a hint that one thing that gets added is that
the PCs of corresponding traces always stay in lock step.)
